---
title: "AMP-AD Annotations"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default-fonts, default, "style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include = FALSE, results = "hide", message = FALSE}
options(htmltools.dir.version = FALSE)
library("synapser")
library("tidyverse")
library("DT")
synLogin()

## ggplot2 theme
theme_set(theme_grey(base_size = 14))
```

# Unique annotations in AMP-AD

```{r get-all-ampad, echo = FALSE, results = "hide", cache = TRUE}
## Query file view of all individual studies + rnaseq reprocessing and rnaseq
## sample swap
dat_fv <- synTableQuery("select * from syn18396001")
dat_full <- as.data.frame(dat_fv)

## Only take publicly released data
dat <- filter(dat_full, benefactorId == "syn2580853")

## Non-annotation columns
fv_cols <- c(
  "ROW_ID",
  "ROW_VERSION",
  "ROW_ETAG",
  "id",
  "name",
  "createdOn",
  "createdBy",
  "etag",
  "type",
  "currentVersion",
  "parentId",
  "benefactorId",
  "projectId",
  "modifiedOn",
  "modifiedBy",
  "dataFileHandleId",
  "concreteType"
)

annot_cols <- setdiff(names(dat), fv_cols)

## How many files? How many unique annotations?
n_files <- nrow(dat)
n_unique <- dat %>%
  select(!!annot_cols) %>%
  distinct() %>%
  nrow()
```

Across all public individual studies, plus the RNA-Seq reprocessing study and
RNA-Seq sample swap study there are:

* **`r n_files`** files
* **`r n_unique`** unique annotation combinations

`r round((n_unique / n_files) * 100)`% of files are uniquely annotated.

--

The AD Cross Species study has many files and lacks the `specimenID` annotation,
so most of these files are not uniquely annotated. Excluding AD Cross Species, we have:

```{r count-distinct-without-ad-cross-species, echo = FALSE}
dat_except_ad_cross <- dat %>%
  filter(!study %in% c("AD_CrossSpecies", "AD_CrossSpecies "))

n_unique_minus_ad_cross <- dat_except_ad_cross %>%
  select(!!annot_cols) %>%
  distinct() %>%
  nrow()
```

* **`r nrow(dat_except_ad_cross)`** files
* **`r n_unique_minus_ad_cross`** unique annotation combinations

`r round((n_unique_minus_ad_cross / nrow(dat_except_ad_cross)) * 100)`% of files
are uniquely annotated.

---

# Which files share annotations? 

```{r distinct-annotation-groups-count, echo = FALSE}

dat_dist_count <- dat_except_ad_cross %>%
  group_by_at(annot_cols) %>%
  summarize(
    n = n(),
    synIDs = list(id),
    parents = list(unique(parentId))
  ) %>%
  arrange(desc(n))

```

```{r plot-distinct-annotation-groups-count, echo = FALSE, dev = "svg", fig.height = 4.5}
ggplot(dat_dist_count, aes(x = n)) +
  geom_histogram(bins = 40) +
  scale_x_log10() +
  labs(
    x = "Number of files with same annotations",
    y = "Count",
    title = "Counts of files sharing the same set of annotations"
  )
```

---

# Which files share annotations? 

```{r distinct-annotation-groups-table, echo = FALSE}
## Show table of number of files with same annotations, preview of files'
## synIDs, and preview (and count) of parent IDs of files
dat_dist_count %>%
  ungroup() %>%
  select(n, synIDs, parents) %>%
  filter(n > 1) %>%
  pmap_dfr(
    function(synIDs, parents, ...) {
      collapsed_ids <- paste(synIDs, collapse = ", ")
      if (nchar(collapsed_ids) > 20) {
        collapsed_ids <- strtrim(collapsed_ids, 20)
        collapsed_ids <- paste0(collapsed_ids, "...")
      }
      collapsed_parents <- paste0(
        "[",
        length(parents),
        "] ",
        paste(parents, collapse = ", ")
      )
      if (nchar(collapsed_parents) > 20) {
        collapsed_parents <- strtrim(collapsed_parents, 20)
        collapsed_parents <- paste0(collapsed_parents, "...")
      }
      bind_cols(
        ...,
        tibble(
          files = collapsed_ids,
          parents = collapsed_parents
        )
      )
    }
  ) %>%
  datatable(options = list(pageLength = 8))
```

---

# RNA Seq Reprocessing

```{r reprocess, echo = FALSE, results = "hide", message = FALSE, cache = TRUE}
reprocess <- dat %>%
  filter(study %in% c("rnaSeqReprocessing", "rnaSeqReprocessing "))

n_files_reprocess <- nrow(reprocess)
n_unique_reprocess <- reprocess %>%
  select(!!annot_cols) %>%
  distinct() %>%
  nrow()
```

The RNA Seq reprocessing study has `r n_files_reprocess` files with 
`r n_unique_reprocess` unique annotation combinations. Below are the groups of
files with repeated annotations:

```{r reprocess-duplicated, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE}
reprocess_duplicated_annots <- reprocess %>%
  group_by_at(annot_cols) %>%
  summarize(n = n(), synIDs = list(id)) %>%
  filter(n > 1) %>%
  arrange(desc(n))

reprocess_grouped_files <- reprocess_duplicated_annots %>%
  ungroup() %>%
  pmap_dfr( 
    function(synIDs, ...) {
      filenames <- map_chr(synIDs, function(y) {
        file <- synGet(y, downloadFile = FALSE)
        file$properties$name
      })
      filenames <- paste(filenames, collapse = ", ")
      bind_cols(..., tibble(files = filenames))
    }
  ) %>%
  select(n, files)
```

```{r show-table-reprocess, echo = FALSE, cache = TRUE}
datatable(
  reprocess_grouped_files,
  fillContainer = FALSE,
  rownames = FALSE,
  options = list(pageLength = 3)
)
```

---

# Some other issues

- Inconsistent annotation values, e.g. spaces at the end:

```{r inconsistent-study-names, echo = FALSE}
grep("\\s$", unique(dat$study), value = TRUE)
```

---

# Some other issues

* Invalid annotation keys

```{r invalid-keys, echo = FALSE}
library("dccvalidator")
check_annotation_keys(select(dat, !!annot_cols))
```

---

# Some other issues

* Invalid annotation values

```{r invalid-values, echo = FALSE}
invalid_values <- check_annotation_values(dat)
```

Too many to show them all, but there are invalid values for 
`r length(invalid_values)` annotations.
